NumPy---Array Programming in Python

### 2020-01-31 body.tex
Rework last paragraph of ecosystem section

SciPy and Matplotlib are tightly coupled with NumPy---in terms of
history, development, and use.
SciPy provides fundamental algorithms for scientific computing
including mathematical, scientific, and engineering routines.
Matplotlib generates publication-ready figures and visualizations.
The combination of NumPy, SciPy, and Matplotlib together with
an advanced interactive environment like IPython \cite{perez2007ipython}
or Jupyter \cite{Kluyver:2016aa}
provides a complete foundation for array-based scientific computing.
The scientific Python ecosystem (Fig.~\ref{fig:ecosystem}) builds on top of
this foundation to provide several, widely used \emph{technique-specific}
libraries\cite{pedregosa2011scikit,vanderwalt2014scikit,SciPyProceedings_11},
that in turn underlay numerous \emph{domain-specific} projects
\cite{astropy:2013,astropy:2018,cock2009biopython,millman2007analysis,sunpy2015,2018EGUGA..2012146H}.
NumPy, as the focal point for the ecosystem of array-aware libraries,
sets documentation standards, and provides array testing infrastructure
and build support for integrating with Fortran code and using specialized
compilers and computing platforms.

@@

Many research groups have designed large,
complex scientific libraries, which add specialized functionality for
solving problems in specific domains.
For example, the \code{eht-imaging} library developed by the Event Horizon Telescope 
collaboration for radio interferometry imaging, analysis, and simulation
relies on many components of the scientific Python ecosystem.
NumPy arrays are used to store and manipulate numerical data at every step
in the processing chain: from raw data through calibration and image 
reconstruction.
SciPy is used to provide tools used in general image processing such as 
filtering and image alignment, while scikit-image, a popular image processing
library that extends SciPy, provides higher-level functionality such as
edge filters and Hough transforms.
Other components of SciPy are used throughout the library, such as the
\code{scipy.optimize} module for handling general optimization tasks.
NetworkX \cite{SciPyProceedings_11}, a popular package for complex
network analysis, is used for analysis of consistency among image
comparisons.
The community-developed astronomy library, Astropy \cite{astropy:2013, astropy:2018},
is used throughout \code{eht-imaging}, primarily to handle standard
astronomical file formats and time/coordinate transformations.
Matplotlib is used for visualizing data throughout the analysis pipeline,
including the generation of the final image of the black hole.

Similar examples can be found in physics, earth science, biology,
and more.
In many fields, the scientific Python ecosystem is the primary
software environment.
The recent rapid growth of data science, machine learning, and
artificial intelligence has dramatically increased scientific
Python use.
NumPy and its ecosystem is now taught in
numerous university courses, bootcamps, and summer schools.
There are multiple community conferences and workshops.
NumPy and its API has become ubiquitous.

>>>>

SciPy and Matplotlib are tightly coupled with NumPy---in terms of
history, development, and use.
SciPy provides fundamental algorithms for scientific computing
including mathematical, scientific, and engineering routines.
Matplotlib generates publication-ready figures and visualizations.
The combination of NumPy, SciPy, and Matplotlib together with
an advanced interactive environment like IPython \cite{perez2007ipython}
or Jupyter \cite{Kluyver:2016aa}
provides a solid foundation for array-based scientific computing.
The scientific Python ecosystem (Fig.~\ref{fig:ecosystem}) builds on top of
this foundation to provide several, widely used \emph{technique-specific}
libraries\cite{pedregosa2011scikit,vanderwalt2014scikit,SciPyProceedings_11},
that in turn underlay numerous \emph{domain-specific} projects
\cite{astropy:2013,astropy:2018,cock2009biopython,millman2007analysis,sunpy2015,2018EGUGA..2012146H}.
NumPy, as the focal point for the ecosystem of array-aware libraries,
sets documentation standards, and provides array testing infrastructure
and build support for integrating with Fortran code and using specialized
compilers and computing platforms.

@@

Many research groups have designed large,
complex scientific libraries, which add specialized functionality for
solving problems in specific domains.
For example, the \code{eht-imaging} library developed by the Event Horizon Telescope 
collaboration for radio interferometry imaging, analysis, and simulation
relies on many components of the scientific Python ecosystem.
NumPy arrays are used to store and manipulate numerical data at every step
in the processing chain: from raw data through calibration and image 
reconstruction.
SciPy is used to provide tools used in general image processing such as 
filtering and image alignment, while scikit-image, an image processing
library that extends SciPy, provides higher-level functionality such as
edge filters and Hough transforms.
Other components of SciPy are used throughout the library, such as the
\code{scipy.optimize} module for handling general optimization tasks.
NetworkX \cite{SciPyProceedings_11}, a package for complex
network analysis, is used for analysis of consistency among image
comparisons.
The community-developed astronomy library, Astropy \cite{astropy:2013, astropy:2018},
is used throughout \code{eht-imaging}, primarily to handle standard
astronomical file formats and time/coordinate transformations.
Matplotlib is used for visualizing data throughout the analysis pipeline,
including for the generation of the final image of the black hole.

The rapid recent growth of data science, machine learning, and
artificial intelligence has further and dramatically boosted usage of
scientific Python.  Examples of its impactful application, such as the
above, now exist in almost every discipline in the natural and social
sciences, and in many fields these tools have become \emph{the
primary} software environment.  NumPy and its ecosystem are commonly
taught in university courses, bootcamps, and summer schools, and are
at the focus of community conferences and workshops worldwide.

NumPy and its API have become truly ubiquitous.
